---
title: "Simulating Quantities of Interest in a Linear Model"
#output: rmarkdown::html_vignette
#vignette: >
#  %\VignetteIndexEntry{linear-model}
#  %\VignetteEngine{knitr::rmarkdown}
#  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=F}
library(tidyverse)
library(stevemisc)
library(modelr)
library(simqi)
library(stevethemes)

set.seed(8675309) # Jenny, I got your number...
```

## A (Brief) Description of the Data

`som_sample` is a subset of [the SOM surveys in Sweden](https://www.gu.se/en/som-institute/the-som-surveys). It is the Swedish corollary to the General Social Survey in the United States and offers a wide-reaching assessment of Swedish public opinion about various aspects of society, politics, and mass media. Deep knowledge about Sweden is not required for this tutorial though some kind of comfort with modeling public opinion data is assumed.

```{r}
som_sample
```

## A Simple Linear Model with No Frills

Let's run a simple linear model regressing a respondent's latent political trust (`lptrust`) on their age (`age`, in years), whether they self-identify as a woman or not (`female`), and their educational attainment (`edu3`) into "low" (1), "medium" (2), or "high" (3) categories. [The codebook](https://svmiller.com/simqi/reference/som_sample.html) offers more information about how the political trust variable is created. Understand, for now, that higher values in `lptrust` = "more" political trust and the variable approximates a standard normal distribution.

Let's use the `lm()` to regress political trust on those three things.

```{r}
M1 <- lm(lptrust ~ age + female + edu3, som_sample)
summary(M1)
```

We find no discernible effect of age on political trust. Partialing out gender and education, young and old are not discernibly different in their latent political trust. The effect of `female` and `edu3` is positive and significant.

What if we wanted to unpack that gender difference further. `data_grid()` in `{modelr}` can help us by creating a hypothetical prediction grid based on the data and the model, generating a hypothetical person with the median age (54) and educational attainment ("medium"). This hypothetical person will differ only in their gender. One is a woman and the other is not.

```{r}
som_sample %>%
    data_grid(.model = M1,
              female = c(0, 1)) -> newdat

newdat
```


`sim_qi()` will take the model (`M1`) and the prediction grid we created above as "new data" and simulate some `nsim` values of the dependent variable (default: 1000) from the model's vector of coefficients and the variance-covariance matrix. We can optionally toggle `return_newdata = TRUE` to help us with post-processing.

```{r}
Sims <- sim_qi(M1, nsim = 100, newdata = newdat, return_newdata = TRUE)
Sims
```

The simulated values of latent political trust are returned as a column called `y` in the `Sims` object we created. From there, we can summarize these simulated values of political trust comparing men and women like this.

```{r}
Sims %>%
    summarize(lwr = quantile(y, .05),
              mean = mean(y),
              upr = quantile(y, .95),
              .by = female)
```

There are certainly fancier summary techniques available, especially visually, but the fundamental takeaway suggests that the mean simulated values of political trust for this "typical woman" is .032 whereas it is -.168 for this "typical man". That's a difference of about .20 on this scale. If there were truly no differences between men and women in their expected values of political trust, the simulations betray that. A 90% interval summarizing the distribution for women has a lower bound of -.011. The upper bound for men -.128. The two do not overlap.

```{r, fig.width=9.5, fig.height = 6}
Sims %>%
    summarize(lwr = quantile(y, .05),
              mean = mean(y),
              upr = quantile(y, .95),
              .by = female) %>%
    mutate(cat = c("Men", "Women")) %>%
    ggplot(.,aes(cat, mean, ymin=lwr, ymax=upr)) +
    theme_steve() +
    geom_pointrange() +
    coord_flip() +
    labs(y = "Simulated Values of Political Trust (with 90% Intervals)",
         x = "",
         title = "Simulated Values of Latent Political Trust, by Men and Women",
         subtitle = "A visual summary better emphasizes what the model communicates, and what the simulations are telling you.",
         caption = "?som_sample in {simqi} (by way of SOM).")
```

## Interactive Effects

What if you were interested in an interactive relationship between age and gender? `{simqi}` is flexible with that as well. First, let's center the age variable by subtracting it from its mean. While it's not strictly necessary for identifying interactive relationships, it is really good practice to make sure there is a naturally occurring zero for all things being interacted. While there is no one with the mean age in the data (i.e. age can only be an integer), we've at least shifted zero from newborns (who won't appear in the data) to the center of the distribution.

```{r}
som_sample %>%
    mutate(c_age = age - mean(age)) -> Data

M2 <- lm(lptrust ~ c_age*female + edu3, Data)
summary(M2)
```

The results of this simple model suggest that 1) higher levels of age coincide with more political trust among those that are not women, 2) women of the average age have higher levels of political trust than non-women, and 3) there is indeed a statistically significant interaction between age and gender.

But what does that "look like"? Here, `sim_qi()` comes to the rescue. Let's create another `newdata` data frame, but this one where we also allow the age variable to take on all unique values (to approximate the full range of age).

```{r}
Data %>%
    data_grid(.model = M2,
              female = c(0, 1),
              c_age = unique(c_age)) -> newdat

newdat
```

Admittedly, getting used to transformed variables like this can be wonky if you're new. So, let's get the actual age equivalents for `c_age` just to have that information handy for post-processing.


```{r}
Data %>%
    distinct(c_age, age) %>% arrange(age) %>%
    left_join(newdat, .) -> newdat

newdat
```

Now, let `sim_qi()` do its thing.


```{r}
Sims <- sim_qi(M2, nsim = 100, newdata = newdat, return_newdata = TRUE)
Sims
```

And summarize accordingly....

```{r, fig.width=9.5, fig.height = 6}
Sims %>%
    mutate(cat = ifelse(female == 1, "Women", "Not Women")) %>%
        summarize(lwr = quantile(y, .05),
              mean = mean(y),
              upr = quantile(y, .95),
              .by = c(cat, age)) %>%
    ggplot(.,aes(age, mean, ymin=lwr, ymax=upr, color=cat, linetype=cat, fill=cat)) +
    theme_steve() +
    geom_line() +
    geom_ribbon(alpha=.2, color='black') +
    scale_x_continuous(breaks = seq(20, 80, by= 10)) +
    labs(title = "Simulated Values of Political Trust, Interacting Gender and Age",
         subtitle = "Interaction terms can be wonky to interpret from the regression output. Simulate them to make sense of them!",
         y = "Simulated Values of Political Trust (with 90% Intervals)",
         x = "",
         color = "", linetype = "", fill = "",
         caption = "?som_sample in {simqi} (by way of SOM).")
```

## Square Terms and the Like

`sim_qi()` works just fine with square terms you might include by way of the `I()` function in base R. Here, let's drop the interactive effect of interest and explore a hypothesis of a curvilinear relationship between age and political trust while partialing out the effects of the gender variable and the education category. You'd specify such a model like this, using the centered age variable we just created.

```{r}
M3 <- lm(lptrust ~ c_age + I(c_age^2) + female + edu3, Data)
summary(M3)
```

The results suggest a potential non-linearity in age. At the least, the square term for age is statistically significant but the other term is not at any conventional threshold. However, what that means for inferential takeaways wouldn't be so clear from the regression output. So, let's use `sim_qi()` to make sense of this. We'll borrow the `newdat` data frame we created above, but subset them to just the women for ease of illustration. Then, let `sim_qi()` do its thing.

```{r}
newdat %>%
    filter(female == 1) -> newdat

Sims <- sim_qi(M3, nsim = 100, newdata = newdat, return_newdata = TRUE)

Sims

```

Now, let's summarize what we just did to make sense of what the regression model is telling us.

```{r, fig.width=9.5, fig.height = 6}
Sims %>%
        summarize(lwr = quantile(y, .05),
              mean = mean(y),
              upr = quantile(y, .95),
              .by = c(age)) %>%
    ggplot(.,aes(age, mean, ymin=lwr, ymax=upr)) +
    theme_steve() +
     geom_line() +
    geom_ribbon(alpha=.2, color='black') +
    scale_x_continuous(breaks = seq(20, 80, by= 10)) +
    labs(title = "Simulated Values of Age and Political Trust for Women",
         subtitle = "We got some indication of non-linearity in the model, but simulation tells us what it 'looks like'.",
         y = "Simulated Values of Political Trust (with 90% Intervals)",
         x = "",
         color = "", linetype = "", fill = "",
         caption = "?som_sample in {simqi} (by way of SOM).")
```
